A Comparison of Parallel Graph Processing Implementations

### 2017-05-15 papers/publication/paper-pollard.tex
mostly results discussion edits

\begin{table}
	\caption{A tabulated sample run of Graphalytics with 32 threads. Just one run per experiment is performed. Runtimes are in seconds. Below the table is an excerpt from the GraphMat log file.}
	\centering
	% cit-Patents: (3 774 768 vertices, 16 518 948 edges)
	% dota-league (61 170 vertices, 50 870 313 edges)
	\begin{tabular}{l|c|c|c|c|c|c}
	\multicolumn{5}{l}{\large{\emph{Graphalytics}}} & \\
	GraphBIG & BFS  & CDLP  &   LCC   &  PR &  SSSP & WCC \\ \hline
	cit-Patents & 0.8 & 11.8 & 15.5  &  4.5 & N/A  & 1.3 \\
	dota-league & 1.1 & 3.9 & 1073.7 &  2.6 & \textbf{3.0} & 1.0 \\
	\multicolumn{1}{l}{}\\
	PowerGraph & BFS  & CDLP  &   LCC   &  PR &  SSSP & WCC \\ \hline
	cit-Patents & 13.8 & 30.1 & 23.9 & 18.8 & N/A & 22.1 \\
	dota-league & 25.6 & 31.2 & 458.1 & 26.7 & 28.9 & 22.9 \\
	% graph500-22 43.0 s 55.6 s 299.8 s 46.4 s N/A    40.5 s
	\multicolumn{1}{l}{}\\
	GraphMat & BFS  & CDLP  &   LCC   &  PR &  SSSP & WCC \\ \hline
	cit-Patents & 7.8 & 20.6 & 11.2 &  8.4 & N/A &  7.0 \\
	dota-league & 2.2 & 19.7 & 150.1 & 2.8 & \textbf{3.8} & 3.1
	\end{tabular} \\
	\vspace{12pt}
	Timing results (for GraphMat SSSP on dota-league)
	\begin{itemize}
		\item Finished file read of dota-league. time: 0.9495
		\item load graph: 3.48167 sec
		\item initialize engine: 3.60012e-05 sec
		\item run algorithm: 0.168526 sec
		\item print output: 0.105846 sec
		\item deinitialize engine: 6.00815e-05 sec
	\end{itemize}
	\label{tab:graphalytics-realworld}
\end{table}

@@

Beyond this plethora of performance data, the Graph Algorithm Platform Benchmark Suite (GAP)~\cite{Beamer:2015:GAPBench}, GraphBIG~\cite{Nai:2015:Graphbig}, and the Graph500~\cite{Murphy:2010:Graph500} consider themselves reference implementations or benchmark suites next to which other implementations can be compared. If no fewer than three software packages call themselves, ``reference implementations,'' which one are we to trust? Our belief is choosing a single reference implementation cannot capture the complexities inherent with graph processing.

@@

We refer to a graph's \emph{scale} when describing the size of the graph. Specifically, a graph with scale $S$ has $2^S$ vertices. For example, many of our experiments were run with $2^{22} = 4,194,304$ vertices and an average of 16 edges per vertex. We measure strong scaling and speedup by varying the number of threads from one to the total number of threads on our research computer, 72.

@@

\subsection{Scalability}
\todo{People want more explanation of this}Figure~\ref{fig:bfs-scaling} shows the scalability and speedup for BFS. We define the strong scaling as $T_1 / (n T_n)$ where $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ is the time for $n$ threads. A linear scalability is when $T_n = T_1/n$ and is the horizontal line on the bottom plot in Fig.~\ref{fig:bfs-scaling}. These plots show in general poor scaling throughout, a challenge facing most parallel graph algorithms. Additionally, $2^{20}$ vertices is a ``small'' graph by today's standards and thus library designers focus on scalability for larger graphs. Another limitation may be the ability for OpenMP to efficiently handle such a large number of threads per machine. GCC version 4.8.5 supports OpenMP version 3.1 but the most recent release of OpenMP is 4.5.
\begin{figure}
	\centering
	\begin{minipage}{\linewidth}
		% trim=left lower right upper
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_speedup20.pdf}
	\end{minipage}
	\begin{minipage}{\linewidth}
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_ss20.pdf}
	\end{minipage}
	\caption{The top figure has both axes logarithmic with an exception at 72 threads for readability. Linear speedup is. GraphBIG dips below 1 because it is slower for $2$ threads than for 1. $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ time for $n$ threads.}
	\label{fig:bfs-scaling}
\end{figure}

@@

One overarching issue with these software packages is the speed at which they change. On one end of the spectrum, the code base of PowerGraph was made closed-source in 2015. Simple fixes such as increasing the maximum number of threads per process and updating dependency URLs are not merged and as a result there is much duplicated effort among more than 400 forks of PowerGraph. On the other end, GraphMat's update to version 2.0 is in general incompatible with version 1.0 and as such has not yet been adopted by Graphalytics nor our framework. One potential solution would be to add each framework to a package management system such as Spack \cite{Gamblin:2015:Spack} which focuses on High Performance Computing.

>>>>

\begin{table}
	\caption{Graphalytics: tabulated sample run times (seconds) with 32 threads. Just one run per experiment is performed. Below the table is an excerpt from the GraphMat log file.}
		\label{tab:graphalytics-realworld}
	\vspace{-10pt}
	\begin{center}
	% cit-Patents: (3 774 768 vertices, 16 518 948 edges)
	% dota-league (61 170 vertices, 50 870 313 edges)
	\begin{tabular}{l|c|c|c|c|c|c}
	%% BN: Commented out because it's redundant with the caption
	%\multicolumn{5}{l}{\large{\emph{Graphalytics}}}  \\
	GraphBIG & BFS  & CDLP  &   LCC   &  PR &  SSSP & WCC \\ \hline
	cit-Patents & 0.8 & 11.8 & 15.5  &  4.5 & N/A  & 1.3 \\
	dota-league & 1.1 & 3.9 & 1073.7 &  2.6 & \textbf{3.0} & 1.0 \\
	\multicolumn{1}{l}{}\\
	PowerGraph & BFS  & CDLP  &   LCC   &  PR &  SSSP & WCC \\ \hline
	cit-Patents & 13.8 & 30.1 & 23.9 & 18.8 & N/A & 22.1 \\
	dota-league & 25.6 & 31.2 & 458.1 & 26.7 & 28.9 & 22.9 \\
	% graph500-22 43.0 s 55.6 s 299.8 s 46.4 s N/A    40.5 s
	\multicolumn{1}{l}{}\\
	GraphMat & BFS  & CDLP  &   LCC   &  PR &  SSSP & WCC \\ \hline
	cit-Patents & 7.8 & 20.6 & 11.2 &  8.4 & N/A &  7.0 \\
	dota-league & 2.2 & 19.7 & 150.1 & 2.8 & \textbf{3.8} & 3.1\\ \hline
	\end{tabular}
	\end{center}
	%\vspace{12pt}
\noindent Timing results (for GraphMat SSSP on dota-league)
	\begin{itemize}
		\item Finished file read of dota-league. time: 0.9495
		\item load graph: 3.48167 sec
		\item initialize engine: 3.60012e-05 sec
		\item run algorithm: 0.168526 sec
		\item print output: 0.105846 sec
		\item deinitialize engine: 6.00815e-05 sec
	\end{itemize}
\end{table}

@@

Beyond this plethora of performance data, the Graph Algorithm Platform Benchmark Suite (GAP)~\cite{Beamer:2015:GAPBench}, GraphBIG~\cite{Nai:2015:Graphbig}, and the Graph500~\cite{Murphy:2010:Graph500} consider themselves reference implementations or benchmark suites with which other implementations can be compared. If no fewer than three software packages call themselves, ``reference implementations,'' which one are we to trust? Our belief is that choosing a single reference implementation cannot capture the complexities inherent in graph processing.

@@

We refer to a graph's \emph{scale} when describing the size of the graph. Specifically, a graph with scale $S$ has $2^S$ vertices. For example, many of our experiments were performed on graphs with $2^{22} = 4,194,304$ vertices and an average of 16 edges per vertex. We measure strong scaling and speedup by varying the number of threads from one to the total number of threads available on our research server, 72.

@@

\subsection{Scalability}
Figures~\ref{fig:bfs-speedup} and~\ref{fig:bfs-efficiency} illustrate the parallel strong scaling of the BFS algorithm in different packages on a Kronecker graph of scale $20$ . The parallel speedup shown in Fig.~\ref{fig:bfs-speedup} is computed as $T_1 / T_n$ where $T_1$ is the sequential time and $T_n$ is the execution time on $n$ threads. 
\begin{figure}[htb]
	\centering
		% trim=left lower right upper
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_speedup20.pdf}
		\caption{Speedup of BFS for a scale-20 graph, black solid line represents ideal speedup. Both axes are logarithmic with an exception at 72 threads for readability. }
	\label{fig:bfs-speedup}
\end{figure}

We define the parallel efficiency as $T_1 / (n T_n)$ where $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ is the time for $n$ threads. Ideal efficiency is defined as $T_n = T_1/n$ and is the horizontal line near the top of Fig.~\ref{fig:bfs-efficiency}. 
\begin{figure}[htb]
	\centering
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_ss20.pdf}
	\caption{Parallel efficiency for a scale-20 graph. GraphBIG dips below 1 because it is slower for $2$ threads than for 1. $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ time with $n$ threads.}
	\label{fig:bfs-efficiency}
\end{figure}

These plots show generally poor scaling for this size problem, a challenge facing most parallel graph algorithms. 
\todo[inline]{If time permits, replace with a larger graph to show scaling.}
Additionally, $2^{20}$ vertices is a relatively ``small'' graph resulting for by today's standards and thus library designers focus on scalability for larger graphs. Another limitation may be the ability for OpenMP to efficiently handle such a large number of threads per machine. GCC version 4.8.5 supports OpenMP version 3.1 but the most recent release of OpenMP is 4.5.

@@

One overarching issue with these software packages is the speed at which they change. On one end of the spectrum, the code base of PowerGraph was made closed-source in 2015. Simple fixes such as increasing the maximum number of threads per process and updating dependency URLs are not merged and as a result there is much duplicated effort among more than 400 forks of PowerGraph. On the other end, GraphMat's update to version 2.0 is in general incompatible with version 1.0 and as such has not yet been adopted by Graphalytics nor our framework. One potential solution would be to add each framework to a package management system such as Spack \cite{Gamblin:2015:Spack} which supports packaging of high performance computing software.
